---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# org.HsSimple.eg.db

<!-- badges: start -->
<!-- badges: end -->

Genome wide annotation for Homo sSimple, primarily based on mapping using Entrez Gene identifiers.

## Installation

To install this package, you need to download latest archive from [release page](https://github.com/bioinformatist/org.HsSimple.eg.db/releases), then run command like:

```r
install.packages("./org.HsSimple.eg.db", repos=NULL)
```

## How can we build this simple database?

Firstly, GO terms in BP and MF ontologies with too many genes were filtered out, for these terms play more general rules which can be represented by other finer terms.

```{r}
library(data.table)
library(ggplot2)
suppressMessages(library(tidyverse))  # For pipe
suppressMessages(library(org.Hs.eg.db))

BP.terms <- as.data.table(AnnotationDbi::select(org.Hs.eg.db, keys = 'BP', columns = c('GO', 'ENTREZID'), keytype = "ONTOLOGY"))
BP.terms[, ONTOLOGY := NULL]
BP.terms <- unique(BP.terms, by = c('GO', 'ENTREZID'))
length(unique(BP.terms$ENTREZID))
BP.terms[, count := .N, by = GO]
setorder(BP.terms, -count)
(BP.terms.uniq <- unique(BP.terms, by = 'GO'))
ggplot() + geom_density(data = BP.terms.uniq, aes(count)) + scale_x_continuous(breaks = seq.int(0, 80, 10), labels = seq.int(0, 80, 10), limits = c(0, 100))
BP.terms <- BP.terms[count < 80]
BP.terms[, count := NULL]
setnames(BP.terms, 3, 'GID')
setcolorder(BP.terms, c('GID', 'GO', 'EVIDENCE'))
go.gene <- dcast(BP.terms, GO ~ GID)
for (j in 2:ncol(go.gene)) set(go.gene, which(!is.na(go.gene[[j]])), j, 1L)
for (j in 2:ncol(go.gene)) set(go.gene, which(is.na(go.gene[[j]])), j, 0L)
go.gene <- go.gene[, lapply(.SD, as.numeric), by = GO]
go.gene <- as.data.frame(go.gene) %>% remove_rownames %>% column_to_rownames(var = "GO")

# For it takes too long to run, I saved the result for reusing
# library(vegan)
# BP.jaccard <- 1 - vegdist(go.gene, method = "jaccard")
load('BP.jaccard.xz')
BP.jaccard <- as.matrix(BP.jaccard)
BP.jaccard[!upper.tri(BP.jaccard)] <- 0
BP.jaccard <- as.data.table(BP.jaccard, keep.rownames = 'GO')
BP.jaccard <- melt(BP.jaccard, id.vars = 'GO')
BP.jaccard <- merge(BP.jaccard, BP.terms.uniq, by.x = 'GO', by.y = 'GO')
BP.jaccard[, `:=`(EVIDENCE = NULL, ENTREZID = NULL)]
setnames(BP.jaccard, c('GO', 'variable', 'count'), c('GO_1', 'GO_2', 'GO_1_count'))
BP.jaccard <- merge(BP.jaccard, BP.terms.uniq, by.x = 'GO_2', by.y = 'GO')
BP.jaccard[, `:=`(EVIDENCE = NULL, ENTREZID = NULL)]
setnames(BP.jaccard, 'count', 'GO_2_count')
# Should mark redundant GO terms as "to be removed" here, or all of them will be kept
BP.jaccard[, remove := ifelse(value > 0.9, ifelse(GO_1_count > GO_2_count, GO_1, GO_2), NA), by = .I]
to.remove <- unique(na.omit(BP.jaccard), by = 'remove')
BP.terms <- BP.terms[setdiff(GO, to.remove[['remove']]), on = 'GO']

CC.terms <- as.data.table(AnnotationDbi::select(org.Hs.eg.db, keys = 'CC', columns = c('GO', 'ENTREZID'), keytype = "ONTOLOGY"))
CC.terms[, ONTOLOGY := NULL]
CC.terms <- unique(CC.terms, by = c('GO', 'ENTREZID'))
length(unique(CC.terms$ENTREZID))
CC.terms[, count := .N, by = GO]
setorder(CC.terms, -count)
(CC.terms.uniq <- unique(CC.terms, by = 'GO'))
ggplot() + geom_density(data = CC.terms.uniq, aes(count)) + scale_x_continuous(limits = c(0, 250))
CC.terms <- CC.terms[count < 175]
CC.terms[, count := NULL]
setnames(CC.terms, 3, 'GID')
setcolorder(CC.terms, c('GID', 'GO', 'EVIDENCE'))
go.gene <- dcast(CC.terms, GO ~ GID)
for (j in 2:ncol(go.gene)) set(go.gene, which(!is.na(go.gene[[j]])), j, 1L)
for (j in 2:ncol(go.gene)) set(go.gene, which(is.na(go.gene[[j]])), j, 0L)
go.gene <- go.gene[, lapply(.SD, as.numeric), by = GO]
go.gene <- as.data.frame(go.gene) %>% remove_rownames %>% column_to_rownames(var = "GO")

# CC.jaccard <- 1 - vegdist(go.gene, method = "jaccard")
load('CC.jaccard.xz')
CC.jaccard <- as.matrix(CC.jaccard)
CC.jaccard[!upper.tri(CC.jaccard)] <- 0
CC.jaccard <- as.data.table(CC.jaccard, keep.rownames = 'GO')
CC.jaccard <- melt(CC.jaccard, id.vars = 'GO')
CC.jaccard <- merge(CC.jaccard, CC.terms.uniq, by.x = 'GO', by.y = 'GO')
CC.jaccard[, `:=`(EVIDENCE = NULL, ENTREZID = NULL)]
setnames(CC.jaccard, c('GO', 'variable', 'count'), c('GO_1', 'GO_2', 'GO_1_count'))
CC.jaccard <- merge(CC.jaccard, CC.terms.uniq, by.x = 'GO_2', by.y = 'GO')
CC.jaccard[, `:=`(EVIDENCE = NULL, ENTREZID = NULL)]
setnames(CC.jaccard, 'count', 'GO_2_count')
CC.jaccard[, remove := ifelse(value > 0.9, ifelse(GO_1_count > GO_2_count, GO_1, GO_2), NA), by = .I]
to.remove <- unique(na.omit(CC.jaccard), by = 'remove')
CC.terms <- CC.terms[setdiff(GO, to.remove[['remove']]), on = 'GO']

MF.terms <- as.data.table(AnnotationDbi::select(org.Hs.eg.db, keys = 'MF', columns = c('GO', 'ENTREZID'), keytype = "ONTOLOGY"))
MF.terms[, ONTOLOGY := NULL]
setnames(MF.terms, 3, 'GID')
setcolorder(MF.terms, c('GID', 'GO', 'EVIDENCE'))

hGO <- rbind(BP.terms, CC.terms, MF.terms)
```

And package directory was built:

```r
AnnotationForge::makeOrgPackage(go = hGO,
               version = "1.0.0",
               maintainer = "Yu Sun <suny226@mail2.sysu.edu.cn>",
               author = "Yu Sun <suny226@mail2.sysu.edu.cn>",
               outputDir = ".",
               tax_id = "9606",
               genus = "Homo",
               species = "sSimple",
               goTable = 'go')
```

We modified the [`zzz.R`](R/zzz.R), first the `dbFileConnect()` function was reloaded for a **read-and-write** connection instead of read-only mode:

```r
dbFileConnect <- function (dbfile)
{
  if (!file.exists(dbfile))
    stop("DB file '", dbfile, "' not found")
  if (.Platform$OS.type == "unix") {
    DBI::dbConnect(RSQLite::SQLite(), dbname = dbfile, cache_size = 64000L,
                   synchronous = "off", flags = RSQLite::SQLITE_RW, vfs = "unix-none")
  }
  else {
    DBI::dbConnect(RSQLite::SQLite(), dbname = dbfile, cache_size = 64000L,
                   synchronous = "off", flags = RSQLite::SQLITE_RW)
  }
}
```

Then, we renamed fields in tables `go_bp`, `go_mf`, `go_cc` and `genes`, for column names provided by `AnnotationForge` package are **incompatible** with functions in `topGO` package.

```r
tryCatch(DBI::dbSendQuery(dbconn, 'ALTER TABLE genes RENAME COLUMN GID to gene_id'), error=function(e) NULL)
lapply(c('go_bp', 'go_cc', 'go_mf'), function(x) tryCatch(DBI::dbSendQuery(dbconn, paste0('ALTER TABLE ', x, ' RENAME COLUMN GO to go_id')), error=function(e) NULL))
```

Finally, the package root directory was packed and published：

```bash
tar cf - org.HsSimple.eg.db | pbzip2 -r > org.HsSimple.eg.db.tar.xz
```

## Contributions

[Yu Sun](https://github.com/bioinformatist/) (maintainer) designed and implemented the database. [Shuang Deng](https://github.com/dengshuang0116/) calculated all Jaccard similarity distance.

## Acknowledgment

We thank Ms. Wenyi Zhou of A.S. Wastons Group for her professional writing advice.